\documentclass[9pt]{article}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,fit}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{hyperref}
\usepackage{balance}
\usepackage[font=small,labelfont=bf]{caption}
\newcommand{\fakeparagraph}[1]{\vspace{3mm}\noindent\textbf{#1.}}
\newcommand{\fakepar}[1]{\fakeparagraph{#1}}
\tikzset{
  block/.style={
    rectangle,
    rounded corners,
    draw,
    align=center,
    minimum height=8mm,
    minimum width=28mm
  },
  decision/.style={
    diamond,
    draw,
    aspect=2,
    align=center,
    inner sep=1pt
  },
  line/.style={
    -{Latex[length=2mm]},
    thick
  }
}
\title{FeBEx: In-Network Deduplication for Federated LoRaWAN Backhaul Using P4 Programmable Switches}

\author{
Savitha Viswanadh Kandala\\
\small{\texttt{
viswanadh@u.nus.edu
}}
% School of Computing, National University of Singapore
\and
Rajashekar Reddy Chinthalapani\\
\small{\texttt{
rajashekar.c@u.nus.edu
}}
% School of Computing, National University of Singapore
}

\date{}

\begin{document}

\maketitle
\vspace{-8mm}
\begin{abstract}
Helium and similar federated LoRaWAN deployments solve the last-mile connectivity problem for IoT devices, but their architecture creates a fundamental inefficiency. When an end device transmits a single uplink packet, multiple geographically distributed hotspots may receive it and independently forward duplicate copies through the backhaul to cloud backends. While this redundancy improves reception reliability and coverage, it also inflates backhaul traffic, increases network cost, and wastes aggregation capacity.

We present FeBEx, a P4-programmable aggregation switch that moves packet deduplication from cloud backends into the network backhaul. FeBEx combines register-based duplicate detection with epoch-driven expiry, longest-prefix-match tenant steering, and packet cloning to preserve proof-of-coverage attribution. We implement FeBEx on a Mininet-BMv2 testbed and evaluate it across deployments with 50--500 devices and 5--50 hotspots. Our results show that FeBEx reduces redundant backhaul traffic by 50--90\% while preserving correct packet delivery, maintaining 100\% multi-tenant isolation, and enabling accurate payment attribution. We further identify practical operating guidelines, including register sizing of at least $4N$ entries and epoch intervals of 5--10 seconds. These results demonstrate that in-network deduplication is a practical and effective approach for reducing redundant backhaul traffic in federated LoRaWAN infrastructures. Find the code and implementation: \url{https://github.com/VIS-WA/FeBEx}.
\end{abstract}
\begingroup
\renewcommand{\thefootnote}{}
\footnotetext{\noindent\textbf{AI Tool Declaration:} We used ChatGPT (GPT-5.4 Thinking) and Claude (Opus 4.6 Extended) to generate ideas, improve wording and paragraph structure, refine technical explanations, assist with coding, and help draft and revise parts of this report, including figure captions and section-level writeups. We are responsible for the content, accuracy, and quality of the submitted work.}
\addtocounter{footnote}{-1}
\endgroup
\input{sections/intro}
% \input{sections/related}
\input{sections/background}
\input{sections/design}
\input{sections/implementation}
\input{sections/evaluation}
% \input{sections/discussion}
\input{sections/conclusion}

\appendix
\input{sections/appendix}

\bibliographystyle{unsrt}
\bibliography{references}

\end{document}